---
title       : Introduction to Kernel Non-Parametric Estimation of Statistical Models
subtitle    : 
author      : Julián D. Arias Londoño
job         : Coursera course - Developing Data Products 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax, bootstrap, quiz, shiny, interactive]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
ext_widgets: {rCharts: [libraries/nvd3]}

--- .class #id 

## Introduction to Non-Parametric Statistical Models

Statistical models are very useful tools in several fields of Engineering and Sciences. Perhaps two of the most commonly used statistical models are: 

<ul>
<li>Multiple Linear Regression or</li>
<li>Multivariate Gaussian Density function based on the Maximum Likelihood criterion. </li>
</ul> 


One of the main drawbacks of these kind of models is the implicit assumption about the nature of the data. <b> Non-parametrics (NP) models </b>, are a class of data-driven statistical models that can be build without making assumptions about the data structure. 

In the following slides, the application of NP models is shown in three different context: 

<ul>
<li>Probability Density Estimation</li>
<li>Regression </li>
<li>Classification </li>
</ul> 

--- .class #id 

## Probabilistic Density Functions

<font size="2"> The most used NP density estimator is the Histogram, which provides a hard estimation of a density function. There exist another kind of NP estimators that can provide smooth estimations in the same context. One of them is the <b>Parzen Window</b> or <b>Kernel</b> estimator. Given a set of samples $\left\lbrace {\bf{x}}_j\right\rbrace _{j=1}^N$, the kernel density estimator is given by: $p({\bf{x}}^*) = \frac{1}{Nh}\sum_{j=1}^N K(({\bf{x}}_j - {\bf{x}}^*)/h)$, where $K(\cdot)$ is the kernel function and $h$ is the bandwidth. </font> 

```{r opts.label = 'shiny'}
library(shiny)
library(slidify)
library(XML)
require(devtools)
slidifyUI(
  sidebarPanel(
    sliderInput("n",label = h4("Number of samples"),
                    min = 10,
                    max = 200,
                    value = 100),
    sliderInput("h",
                       label = h4("Bandwidth Kernel estimatior"),
                       min = 0.01,
                       max = 2,
                       value = 0.5)
    ),
  mainPanel(
    tabsetPanel(
             tabPanel("Histogram", plotOutput("distPlot"),height = "200px",width = "400px"),
             tabPanel("Kernel function", "The kernel function used is: Radial Basis Function")
             )
    )
  )
```


--- .class #id 

## Kernel Regression

<font size="4"> In the same way as before, we can use NP estimators for regression problems. For the sake of comparison, here we are going to use the Nadaraya-Watson kernel regression and a k-Nearest Neighbor estimation.  </font> 

```{r opts.label = 'shiny'}
library(shiny)
library(slidify)
library(XML)
require("proxy")
require(devtools)
slidifyUI(
  sidebarPanel(
    sliderInput("n2",label = h4("Number of samples"),
                    min = 10,
                    max = 200,
                    value = 100),
    sliderInput("h2",
                       label = h4("Bandwidth Kernel estimatior"),
                       min = 0.01,
                       max = 2,
                       value = 0.5),
    sliderInput("k",
                       label = h4("Number of Nearest Neighbors"),
                       min = 1,
                       max = 20,
                       value = 5)
    ),
  mainPanel(
    tabsetPanel(
             tabPanel("Kernel Regressor", plotOutput("distPlot2")),
             tabPanel("K-NN Regressor", plotOutput("distPlot3"))
             )
    )
  )
```

--- .class #id 

## Kernel Decision boundary

```{r opts.label = 'shiny'}
require(rCharts)
slidifyUI(
  sidebarPanel(
    selectInput('Features', 'Choose Features', c('SepalLength~SepalWidth', 'SepalLength~PetalWidth', 'SepalLength~PetalLength', 'PetalLength~PetalWidth', 'SepalWidth~PetalLength', 'SepalWidth~PetalWidth')),
    sliderInput("h3",
                       label = h4("Bandwidth Kernel estimatior"),
                       min = 0.01,
                       max = 2,
                       value = 0.5)
    ),
  mainPanel(
    tabsetPanel(
             tabPanel("Scatter Plot", tags$div(id = 'chart2', class='shiny-html-output nvd3 rChart')),
             tabPanel("Decision Boundary", plotOutput("distPlot5"))
             )
    )
  )
```
